A Spiking Independent Accumulator Model for Winner-Take-All

### 2017-01-22 doc/paper/paper.tex
Finishing methods; edit consistency/simplicity throughout

\begin{abstract}
The winner-take-all (WTA) mechanism is an important component of many cognitive models.
\todo{Flesh this out a bit}
We first provide a novel spiking implementation of the well-known leaky competing accumulator (LCA) model, by mapping its dynamics onto a population-level representation.
We then propose a two-layer spiking independent accumulator (IA) model, and compare its performance against the LCA network on a variety of WTA benchmarks.
Our findings suggest that while the LCA network can rapidly adapt to new winners, the IA network is better suited for stable decision making in the presence of noise.

@@

\section{Introduction}
Winner-take-all (WTA) mechanisms are often employed in cognitive models~\cite<e.g.,>{oreilly1998}.
A WTA mechanism receives a $d$-dimensional input of utility values for $d$ different choices.
The output is supposed to be larger than zero for the dimension with highest utility and zero for all other inputs.
\todo{Say some obvious things, such as the role of WTA in decision making?}

A large body of literature exists examining the optimality of WTA mechanisms and their consistency with neurobiological and psychological data~\cite<e.g.,>{bogacz2006,gold2007,smith2004}.
Here, we investigate the suitability of two different WTA mechanisms in the context of large-scale cognitive modelling.
In particular, we map each mechanism onto a biologically plausible network of spiking neurons, and then compare them using a set of benchmarks that
are normative in nature.
The first mechanism is an implementation of the leaky, competing accumulator model by~\citeA{usher2001}.
It has been widely used, for example in versions of the Temporal Context Model~\cite{sederberg2008}, and some of our models~\cite<e.g.,>{kajic2017}. \todo{How does the latter implementation differ from the implementation here?}
The second mechanism is our proposed independent accumulator (IA) model, which involves a second thresholding layer that is recurrently connected to a primary integrating layer.
For both mechanisms we are interested in decision-making tasks involving more than two choices ($d > 2$). \todo{Misplaced sentence.}

To implement these models, we use the Neural Engineering Framework~\cite<NEF;>{eliasmith2003} to map some prescribed dynamics onto populations of spiking neurons.
The NEF has been used for a wide variety of models, including the largest functioning simulation of the human brain~\cite{eliasmith2012}.
We first provide a short introduction to the NEF, and then describe the implementation of the two WTA mechanisms.
We then describe our benchmarks, and discuss some resulting implications for cognitive modelling.

@@

\todo{Bold font for vectors}

\subsection{The Neural Engineering Framework}
We employ the Neural Engineering Framework~\cite<NEF;>{eliasmith2003} which allows us to implement equations describing the model dynamics in spiking neural networks based on three principles: \emph{representation}, 
\emph{transformation}, and \emph{dynamics}.

\subsubsection{Principle 1: Representation}
The representation of a value $x(t)$ with a population of neurons is defined by the encoding and decoding.
The encoding converting the value $x$ into a spike train or instantaneous firing rate $a_i(t)$ for neuron $i$ is given by
\begin{equation}
    a_i(t) = G_i\left[\alpha_i x(t) + J_i^{\mathrm{bias}}\right]
\end{equation}
where $\alpha_i$ is a gain factor, $J_i^{\mathrm{bias}}$ a bias current and $G_i$ is the neuron non-linearity.
Here, we use spiking, leaky integrate-and-fire neurons as $G_i$.
Decoding weights $d_i$ are used to decode a value $\hat x(t)$ from such a population of neurons with
\begin{equation}
    \hat x(t) = \sum_i d_i \left[a_i(t) * h(t)\right]
\end{equation}
where $h(t) = \tau_{\mathrm{s}}^{-1}\exp(t/\tau_{\mathrm{s}})$ is an exponential synaptic filter with time constant $\tau_{\mathrm{s}}$.
The decoding weights are obtained with a least-squares optimization of the error $E_x = \abs{x     - \hat{x}}$.
For the transmission of a value from one population to another, the connection weights are given by
\begin{equation}
    W_{ij} = \alpha_j d_i \text{.}
\end{equation}

\subsubsection{Principle 2: Transformation}
By finding alternate decoding weights $d^f_i$ with the error given by $E_{f(x)} 
= \abs{f(x) - \hat x}$ arbitrary linear and non-linear functions $f(x)$ can be approximated in the connections between neural populations.

\subsubsection{Principle 3: Dynamics}
TODO \subsection{Leaky, competing accumulator model}
We implement the widely-used leaky, competing accumulator (LCA) model proposed by~\citeA{usher2001}.
The dynamics for the state variables $x_i,\ i < N$, where $N$ is the number of choices, are given by:
\begin{equation} \label{eqn:usher-mcclelland}
    \begin{split}
        \frac{{\partial x}_i}{\partial t} &= \left(\rho_i - kx_i - \beta \sum_{j \neq i} x_j\right) \frac{1}{\tau} \\
        x_i &\rightarrow \max(x_i, 0) ,
    \end{split}
\end{equation}
where $\rho_i$ is the external input, $k$ is the leak rate, $\beta$ the lateral inhibition, and $\tau$ the time constant.
We set $k = \beta = 1$, since this will guarantee that the winning state converges to the value of its input, while the losing states converge to zero.
Other choices of $k$ merely alter the effective $\tau$ and the effective gain on the input, while other choices of $\beta$ will produce unwanted behaviour (see supplementary analysis).
Equation~\ref{eqn:usher-mcclelland} can easily be implemented with the NEF by using one population of neurons for each $x_i$.
We believe this implementation to be novel as it it does not treat the $x_i$ as neural firing rates, rather the neural firing rates are weighted by the decoding weights to precisely implement the stated dynamics.
As such the model equations do not impose restrictions on the neuron model or requires to lump the neurons into a single population firing rate.
This allows us to use biologically more realistic neurons while at the same time adhering to the dynamics prescribed by the model.
Figure~\ref{fig:usher-mcclelland} shows an example output.
\begin{figure}
    \centering
    \includegraphics{figures/usher-mcclelland}
    \caption{Example time course of the state variables in the leaky, competing 
        accumulator model with three dimensions. The vector of inputs is given 
        by $(0.8, 0.7, 0.6)$.}\label{fig:usher-mcclelland}
\end{figure}

\subsection{Independent accumulator model}
The other WTA mechanism that we investigate is our proposed independent accumulator (IA) model.
We use the term `independent' to refer to the fact that there are no direct interactions between each accumulator.
To enable competition, we add a second thresholding layer that will project back to self-excite and mutually-inhibit the first layer.
We now provide the details of each layer~(Fig.~\todo{network diagram}).
The first layer consists of a separate integrator population (i.e., accumulator) for each input.
The second layer consists of independent, non-recurrent populations that receive input from the first layer in a one-to-one fashion.
From the second layer, we decode the shifted Heaviside function $\bar{x}_{2,i} = \Theta(x_{2,i} - \vartheta)$ where $\vartheta = 0.8$ is a threshold value controlling how much evidence needs to be accumulated to produce an output.
The Heaviside decoded output of layer 2 projects back to layer 1 to add <<<<<<< HEAD
$\bar{x}_{2,i} - \bar{\beta} \sum_{j \neq i} \bar{x}_{2,j}$ to the input of $x_{1,i}$.
For $\bar{\beta} = 2$ (see supplementary analysis), the correct choice will self-excite while all other state variables will be suppressed once one of the accumulators reaches the threshold $\vartheta$.
The dynamics are described by the following equations:
\begin{equation}
    \begin{split}
        \frac{{\partial x}_{1,i}}{\partial t} &= \lambda \rho_i \frac{1}{\tau_1} + \left( 
            \bar{x}_{2,i} - \bar{\beta} \sum_{j \neq i} \bar{x}_{2,j} \right) \frac{1}{\tau_2} \\
        x_{1,i} &\rightarrow \max(x_{1,i}, 0) .
    \end{split}
\end{equation}

Instead of controlling the threshold $\vartheta$, in the context of the NEF, it is more convenient to keep $\vartheta$ fixed and scale the input with the $\lambda$ parameter.
Mathematically both approaches are equivalent.
We set $\lambda = 1$ where not noted otherwise.
As the LCA model does not have a threshold parameter, we do not introduce the input scaling $\lambda$ in that model.
Figure~\ref{fig:indacc} shows an example of the model dynamics.
\begin{figure}
    \centering
    \includegraphics{figures/indacc}
    \caption{Example time course of the state variables $x_{1,i}$ and decoded 
        output $\bar{x}_{2,i}$ in the independent accumulator model with three 
        dimensions. The vector of inputs is given by (0.8, 0.7, 0.6).
        \todo{Put next to (or below) first figure.}
    }\label{fig:indacc}
\end{figure}

\subsection{Benchmarks}
To test and compare the two WTA mechanisms we provide an input of $\rho_i 
= u - s \delta_{1i} + \eta_i$ where $u$ is the strength of the strongest input, 
$s$ the target separation to all other inputs, $\delta$ the Kronecker delta, and $\eta_i$ Gaussian white noise with standard deviation $\sigma$.
Thus, without loss of generality, the first state variable receives the strongest input and all other state variables receive input that is smaller by $s$.
We chose this input as for $\rho_i \rightarrow 0$ the influence of input $\rho_i$ will decrease and the effective number of choices to decide between would decrease As $s \rightarrow 0$ the problem gets more difficult.
We will use $u = 1$ as long as not indicated otherwise and set the number of dimension to $d = 10$.
In both WTA models we use 200 neurons for each dimensions.
In the IA network this are split as 150 neurons for each layer 1 population and 50 neurons for each layer 2 population.
Further parameters are summarized in Table~\ref{tbl:params}.
\begin{table}
    \caption{Summary of parameter values.}\label{tbl:params}
    \begin{tabular}{ll}
        LCA time constant & $\tau = \SI{0.1}{\second}$ \\
        LCA recurrent parameters & $k = \beta = 1$ \\
        IA accumulation time constant & $\tau_1 = \SI{0.1}{\second}$ \\
        IA feed-back time constant & $\tau_2 = \SI{0.1}{\second}$ \\
        IA threshold & $\vartheta = 0.8$ \\
        IA recurrent parameters & $\lambda = 1, \bar{\beta} = 2$ \\
        Recurrent synaptic time constant & $\tau_{\mathrm{s}} 
        = \SI{0.1}{\second}$ \\
        Feed-forward synaptic time constant & $\tau_{\mathrm{s}} 
        = \SI{0.005}{\second}$ \\
        Output decoding synaptic time constant & $\tau_{\mathrm{s}} 
        = \SI{0.01}{\second}$
    \end{tabular}
\end{table}

To evaluate the two mechanisms on this input, we use a number of separate metrics.
First we determine whether the model is able to form a clear decision at all within a second of simulated time.
To count as a clear decision, at least one output needs to stay above 0.15 for the time interval
$(\SI{1}{\second}, \SI{2}{\second}]$. % chktex 9
This lower bound was chosen to ensure that noise on a zero output is not mistaken as a non-zero output.
Furthermore, we require that the largest output will be the largest output for all of the time interval or in other words the decision does not change during that time interval.
This does not take into account whether the winning output actually corresponds to the strongest input, but sometimes it can be more desirable to produce a clear decision than actually finding the strongest input in all instances.
In other situations, though, the correctness of the decision might be of higher importance.
We consider a trial correct if the model forms a clear decision and the largest output corresponds to the largest input.
All remaining metrics will be calculated on the subset of trials with a clear decision.
The time it takes to form a decision can be relevant too.
We define the decision time as how long it takes to fulfill the conditions of a clear decision.
These metrics cover the basic performance of the WTA mechanisms.
When using a WTA mechanism within larger cognitive models, the exact magnitude of the outputs can be important.
Ideally, the output would be 1 for the winner and zero for all other outputs.\footnote{Technically, the winner output could be scaled     to any magnitude.
We use 1 here as it is most convenient to work with.} The LCA dynamics, however, prescribe an output of $u$ for the winner with $u$ being the largest input.
We define the \emph{winner error} $E_1$ as the difference of the winner output from $u$ for the LCA model, as the difference from 1 for the IA model and, for both models, the \emph{runner-up error} $E_2$ as the difference of second largest output from 0 (negative values treated as 0).
We use the average output values of the time   interval
 $(\SI{1}{\second}, \SI{2}{\second}]$ % chktex 9
to calculate these error values as the values fluctuate over time.
These two metrics only look at the final output averaged over time, but the model can produce transient outputs before the final decision is achieved.
Within the context of a larger model, this can be a problem as the transient output could already be interpreted as a decision.
Thus, we define $E_{2,\max}$ 
as the highest output of a non-winning output during the whole simulation.

\section{Results}
We find that the ability to reach a clear decision depends on the magnitude of the highest input for both networks.
This is not surprising.
In the LCA model the output magnitude will be determined by the input magnitude of the largest input.
As such the largest input needs to be above 0.15 to fulfill the requirements.
For the independent accumulator model the input has to be large enough to reach the integration threshold within the time limit.
By increasing the input scaling $\lambda$ or the time limit one could get a clear decision for smaller inputs within the time limit.
We further find that the IA model is independent of the input noise and target separation for all tested parameters.
This is not the case for the LCA model as shown in Figure~\ref{fig:decisions} 
where the fraction of trials without a clear decision increases with noise and smaller target separation.
\begin{figure}
    \centering
    \includegraphics{figures/decisions}
    \caption{Fraction of trials with a clear decision for different input noise 
        standard deviations $\sigma$ and target separations $s$ for the LCA 
        model. Error bars denote bootstrapped 95\% confidence intervals.  The 
        grey horizontal line shows the optimum.}\label{fig:decisions}
\end{figure}

However, when considering the correctness of the decisions (see Fig.~\ref{fig:correct}) the LCA model is doing better.
We can alleviate that difference to some degree by reducing the input scaling to $\lambda = 0.2$, but perform still worse for some combinations of input noise and target separation when staying below the time limit.
This in some part due to a systematic bias for certain dimensions introduced in the second layer of the IA model.
In the NEF the neuron gain and bias parameters are usually picked from random distributions.
As such the layer 2 populations will become active at slightly different thresholds.
By using identical layer 2 populations, i.e.~identical gain and bias parameters for each population, this bias can be eliminated and the IA model performs better than the LCA model in most instances.
\begin{figure*}
    \centering
    \includegraphics{figures/correct}
    \caption{Fraction of correct trials for different input noise standard 
        deviations $\sigma$, target separations $s$, and models.  Error bars 
        denote bootstrapped 95\% confidence intervals. The grey horizontal line 
        shows the optimum.}\label{fig:correct}
\end{figure*}

The time required to reach a decision in the LCA model depends mostly on the target separation and amount of noise on the input (see Fig.~\ref{fig:time}), 
the magnitude of the largest input is of minor influence.
Also, the confidence intervals indicate a larger variance in the decision times with less target separation and more input noise.
In contrast to that, in the IA model the scaled magnitude of the largest input is the most important factor.
Depending on this magnitude the network can either be faster or slower than the LCA network, but it will need more time to achieve the same fraction of correct responses.
There is also a slight influence of target separation and input noise with an interaction of these two parameters.
\begin{figure*}
    \centering
    \includegraphics{figures/time}
    \caption{Mean decision times for different input noise standard deviations 
        $\sigma$ and target separations $s$ for the LCA and IA model. For the IA 
        model $u = 0.2$ is shown which shows the interaction of target 
        separation and input noise most clearly. Error bars denote bootstrapped 
        95\% confidence intervals.  The grey horizontal lines show the 
        optimum.}\label{fig:time}
\end{figure*}

The output of the IA model matches almost perfectly the specifications of producing an output of 1 for the winner and 0 for all other dimensions.
This is not the case for the LCA model (see Fig.~\ref{fig:error}).
For small target separations the output of the winner will stay below $u$ and this will get worse with increased noise.
Furthermore, the LCA model will have residual outputs of non-winning choices under noisy conditions and with small target separations.
\begin{figure*}
    \centering
    \includegraphics{figures/error}
    \caption{Error on outputs for different input noise standard deviations 
        $\sigma$ and target separations $s$ for the LCA model.  Error bars 
        denote bootstrapped 95\% confidence intervals. The grey horizontal lines 
        show the optimum.}\label{fig:error}
\end{figure*}

Finally, looking at the transient responses both models might produce outputs of non-winning choices (see Fig.~\ref{fig:transient}).
For the LCA model these increase with noise on the input and, for small target separation, can approach the final output of the winning dimensions.
For the IA model, transient outputs are smaller in noisy conditions, but can be higher than for the LCA model in less noisy conditions.
The magnitude of such transient responses is reduced by decreasing the input scaling.
\begin{figure*}
    \centering
    \includegraphics{figures/transient}
    \caption{Highest transient response of non-winning dimensions for different 
        input noise standard deviations $\sigma$ and target separations $s$ for 
        the LCA and IA model. Error bars denote bootstrapped 95\% confidence 
        intervals. The grey horizontal lines show the 
        optimum.}\label{fig:transient}
\end{figure*}

>>>>

\begin{abstract}
The winner-take-all (WTA) mechanism is an important component of many cognitive models.
\todo{Flesh this out a bit}
We provide a novel spiking implementation of the well-known leaky, competing accumulator (LCA) model, by mapping the dynamics onto a population-level representation.
We then propose a two-layer spiking independent accumulator (IA) model, and compare its performance against the LCA network on a variety of WTA benchmarks.
Our findings suggest that while the LCA network can rapidly adapt to new winners, the IA network is better suited for stable decision making in the presence of noise.

@@

\section{Introduction}

The winner-take-all (WTA) mechanism is a component that selects the largest value among a number of possible inputs.
More specifically, the input is a $d$-dimensional vector corresponding to the non-negative utility of $d$ different choices, and the desired output is positive for the dimension with highest utility (i.e., the ``winner'') and zero for all others.
This mechanism is regularly employed in cognitive models of decision making, which must select the most important action to take at any given moment in time~\cite<e.g.,>{oreilly1998}. 
\todo{Cite survey or cite other examples}

A large body of literature exists examining the optimality of WTA mechanisms and their consistency with neurobiological and psychological data~\cite<e.g.,>{bogacz2006,gold2007,smith2004}.
Here, we investigate the suitability of two different WTA mechanisms in the context of large-scale cognitive modelling.
In particular, we map each mechanism onto a biologically plausible network of spiking neurons, and then compare them using a set of benchmarks that are normative in nature.
The first mechanism is an implementation of the {\it leaky, competing accumulator} (LCA) model by~\citeA{usher2001}.
It has been widely used, for example in versions of the Temporal Context Model~\cite{sederberg2008}, and in some of our models~\cite<e.g.,>{kajic2017}.
\todo{Add sentence somewhere differentiating our implementation from the latter implementation}
The second mechanism is our proposed {\it independent accumulator} (IA) model, which involves a second thresholding layer that is recurrently connected to a primary integrating layer.
For both mechanisms we are interested in decision-making tasks involving more than two choices ($d > 2$).
\todo{Misplaced sentence / unimportant here}

To implement each model, we use the Neural Engineering Framework~\cite<NEF;>{eliasmith2003} to map their dynamics onto populations of spiking neurons.
The NEF has been used for a wide variety of models, including the largest functioning simulation of the human brain~\cite{eliasmith2012}.
We provide a short introduction to the NEF, describe our implementation of the two WTA mechanisms, present our benchmarks, and finally discuss some resulting implications for cognitive modelling.

@@

\subsection{The Neural Engineering Framework}
The Neural Engineering Framework~\cite<NEF;>{eliasmith2003} is a method of mapping a cognitive model, described using mathematical equations, onto a spiking neural network.
We now summarize the relevant aspects of this framework, ordered by its three principles: \emph{representation}, \emph{transformation}, and \emph{dynamics}.

\subsubsection{Principle 1: Representation}
We define the representation of a scalar value $x(t)$ by an encoding and decoding with respect to some population of neurons.
The encoding of $x(t)$ into a spike train $a_i(t)$ for neuron $i$ is given by:
\begin{equation}
    a_i(t) = G_i\left[\alpha_i x(t) + J_i^{\mathrm{bias}}\right] ,
\end{equation}
where $\alpha_i$ is a gain factor, $J_i^{\mathrm{bias}}$ a bias current, and $G_i \left[ \cdot \right]$ is the neural nonlinearity.
Here, we use spiking, leaky integrate-and-fire (LIF) neurons for $G_i \left[ \cdot \right]$.
Decoding weights $d_i$ are then used to approximate the represented value $\hat x(t)$ from the activity of the population of neurons by:
\begin{equation}
    \hat x(t) = \sum_i d_i \left[(a_i * h)(t)\right] ,
\end{equation}
where $h(t) = \tau_{\mathrm{s}}^{-1}\exp(t/\tau_{\mathrm{s}})$ is an exponentially decaying synaptic filter with time-constant $\tau_{\mathrm{s}}$, and $\ast$ is the convolution operator.
The decoding weights are obtained by least-squares optimization of the error $E_x = \abs{x     - \hat{x}}$.
For the transmission of a value from one population to another, the connection weights are given by:
\begin{equation}
    W_{ij} = \alpha_j d_i \text{.}
\end{equation}

\subsubsection{Principle 2: Transformation}
By finding alternate decoding weights $d^f_i$ with the error given by $E_{f(x)} 
= \abs{f(x) - \hat x}$ arbitrary linear and nonlinear functions $f(x)$ can be approximated in the connections between neural populations.

\subsubsection{Principle 3: Dynamics}
Given some nonlinear dynamics for the state variable $x(t)$:
\begin{equation} \label{eqn:nonlinear-system}
    \frac{\partial x}{\partial t} = g(x) ,
\end{equation}
we can map these dynamics onto a recurrent transformation, by harnessing the synaptic filter used in Principle 1.
In particular, for the exponentially decaying $h(t)$ we may apply Principle 2 to the recurrent transformation $f(x) = \tau_s g(x) + x$ to ensure that $x(t)$ obeys (\ref{eqn:nonlinear-system}).

\subsection{Leaky, competing accumulator model}
We implement the widely-used leaky, competing accumulator (LCA) model proposed by~\citeA{usher2001}.
The dynamics for each state variable $x_i(t),\ i < d$, where $d$ is the number of choices, are given by~(see Fig.~\ref{fig:usher-mcclelland}):
\begin{equation} \label{eqn:usher-mcclelland}
    \begin{split}
        \frac{{\partial x}_i}{\partial t} = \left(\rho_i - kx_i - \beta \sum_{j \neq i} x_j\right) \frac{1}{\tau}, \quad x_i &\ge 0,
    \end{split}
\end{equation}
where $\rho_i$ is each external input, $k$ is the leak rate, $\beta$ the lateral inhibition, and $\tau$ the time-constant.
This model essentially integrates each input $\rho_i$ with a leak term ($- kx_i$), minus competition from every other variable ($\beta \sum_{j \neq i} x_j$).
Supposing $\rho_i > \rho_j$ for all $j \ne i$, a WTA mechanism should indicate that $i$ is the winning choice.
Setting $k = \beta = 1$ will guarantee that the winning state $x_i$ converges to the value of the largest input $\rho_i$, while the losing states $x_j$ ($j \ne i$) converge to zero.
Other choices of $k$ merely alter the effective $\tau$ and the effective gain on the input, while other choices of $\beta$ will produce unwanted behaviour (see supplementary analysis).

We implement (\ref{eqn:usher-mcclelland}) with the NEF by using one population of neurons for each $x_i$, and applying Principle~3 to each population.
By appropriately selecting the gain and bias parameters from Principle~1, we ensure that each state variable is rectified ($x_i \ge 0$) as required.
We believe this implementation is novel as it it does not interpret each $x_i$ as a distinct neural firing rate, but rather as a population-level representation distributed across any number of spiking neurons.
In effect, heterogeneous spike trains are weighted by optimal decoding weights to precisely implement the stated dynamics.
%As such, the model equations do not impose restrictions on the neuron model or requires to lump the neurons into a single population firing rate.
This allows us to attain greater biological realism without altering the dynamics prescribed by (\ref{eqn:usher-mcclelland}).

\begin{figure}
    \centering
    \includegraphics{figures/usher-mcclelland}
    \caption{ \label{fig:usher-mcclelland}
        Example time course of the state variables in the LCA network with three choices.
        The vector of inputs is given by $(0.8, 0.7, 0.6)$.
    }
\end{figure}

\subsection{Independent accumulator model}
The other WTA mechanism that we investigate is our proposed independent accumulator (IA) model.
We use the term `independent' to refer to the fact that there are no direct interactions between each accumulator.
To enable competition, we add a second thresholding layer that will project back to self-excite and mutually-inhibit the first layer.
We now provide the details of each layer.
\todo{Network diagram?}

The first layer consists of a separate integrator population (i.e., accumulator) for each state variable $x_i(t), \ i < d$.
The second layer consists of independent, non-recurrent, populations that receive input from the first layer in a one-to-one fashion.
From the second layer, we decode the function $\bar{x}_i := \Theta(x_i - \vartheta)$ where $\Theta$ is the Heaviside function, and $\vartheta = 0.8$ is a fixed threshold value that determines how much evidence needs to be accumulated to produce an output.
The Heaviside decoded output of layer 2 projects back to layer 1 to add $\bar{x}_i - \bar{\beta} \sum_{j \neq i} \bar{x}_j$ to the input of $x_i$.
Since intuitively the largest input will accumulate the fastest, once this reaches the threshold $\vartheta$ it will self-excite and inhibit all other state variables.
Fixing $\bar{\beta} = 2$ ensures that the losing state variables will go to zero (see supplementary analysis).
This is summarized more precisely by the following dynamics (see Fig.~\ref{fig:indacc}):
\begin{equation}
    \begin{split}
        \frac{{\partial x}_i}{\partial t} = \lambda \rho_i \frac{1}{\tau_1} + \left( 
            \bar{x}_i - \bar{\beta} \sum_{j \neq i} \bar{x}_j \right) \frac{1}{\tau_2} , \quad x_i &\ge 0 .
    \end{split}
\end{equation}
We set $\lambda = 1$ to be consistent with the LCA model.
% should we mention that we use Principle 3 to implement each integrator
% should we say anything about how the thresholding population works?

%Instead of controlling the threshold $\vartheta$, it is more convenient to keep $\vartheta$ fixed and scale the input with the $\lambda$ parameter -- mathematically this is equivalent.
% ^ this isn't quit equivalent. you also need to multiply the Heaviside by lambda in order to keep the feedback having the same relative magnitude
% it would be equivalent to scale \tau_1 inversely with \lambda
% also I think this point isn't too important?
%As the LCA model does not have a threshold parameter, we do not introduce the input scaling $\lambda$ in that model.
\begin{figure}
    \centering
    \includegraphics{figures/indacc}
    \caption{ \label{fig:indacc}
        Example time course of the state variables $x_i$ and decoded output $\bar{x}_i$ in the IA network with three choices.
        The vector of inputs is given by (0.8, 0.7, 0.6).
        \todo{Put next to (or below) first figure.}
    }
\end{figure}

\subsection{Benchmarks}
To test and compare the two WTA mechanisms we provide an input of $\rho_i = u - s(1 - \delta_{1i}) + \eta_i$, where $u$ is the strength of the strongest input, 
$s > 0$ the target separation to all other inputs, $\delta$ the Kronecker delta, and $\eta_i$ Gaussian (white) noise with standard deviation $\sigma$.
Thus, without loss of generality, the first state variable receives the strongest input $u$ plus noise, and all other state variables receive a noisy input that is smaller by $s$.
We chose this input as for $\rho_i \rightarrow 0$ the influence of input $\rho_i$ will decrease and the effective number of choices to decide between would decrease.
\todo{I don't understand the previous sentence}
As $s \rightarrow 0$ the problem becomes more difficult.
We will use $u = 1$ unless indicated otherwise, and set the number of choices to $d = 10$.
In both WTA models we use 200 neurons per choice.
In the IA network this is split into 150 neurons for each layer 1 population, and 50 neurons for each layer 2 population.
All remaining network parameters are summarized in Table~\ref{tbl:params}.
\begin{table}
    \caption{Summary of parameter values.}\label{tbl:params}
    \begin{tabular}{ll}
        LCA time-constant & $\tau = \SI{0.1}{\second}$ \\
        LCA recurrent parameters & $k = \beta = 1$ \\
        IA accumulation time-constant & $\tau_1 = \SI{0.1}{\second}$ \\
        IA feedback time-constant & $\tau_2 = \SI{0.1}{\second}$ \\
        IA threshold & $\vartheta = 0.8$ \\
        IA recurrent parameters & $\lambda = 1, \bar{\beta} = 2$ \\
        Recurrent synaptic time-constant & $\tau_{\mathrm{s}} 
        = \SI{0.1}{\second}$ \\
        Feed-forward synaptic time-constant & $\tau_{\mathrm{s}} 
        = \SI{0.005}{\second}$ \\
        Output decoding synaptic time-constant & $\tau_{\mathrm{s}} 
        = \SI{0.01}{\second}$
    \end{tabular}
\end{table}

To evaluate the two mechanisms on this input, we use a number of separate metrics.
First, we determine whether the model is able to form a clear decision at all within a second of simulated time.
To be counted as `clear', at least one output must remain above 0.15 across the time interval $(\SI{1}{\second}, \SI{2}{\second}]$. % chktex 9
This lower bound of 0.15 was chosen to ensure that noise on a zero output is not mistaken for a non-zero output.
Note this requires that the decision does not change throughout the time interval.
This does not take into account whether the winning output actually corresponds to the strongest input, but for some models it is more desirable to produce a clear incorrect decision than an unstable incorrect decision.
In other situations, though, the correctness of the decision may be of higher importance.
Thus, we consider a trial `correct' if the model forms a clear decision, and the largest output corresponds to the true largest input.

We now describe a number of metrics that will be calculated on the subset of trials that contain a clear decision.
% The time it takes to form a decision can be relevant too.
It is important to consider the speed at which the network can make decisions.
We therefore define the `decision time' as the length of time it takes to fulfil the conditions of a clear decision.
%These metrics cover the basic performance of the WTA mechanisms.
Likewise, when using a WTA mechanism within larger cognitive models, the exact magnitude of the outputs can be important.
Ideally, the output would be one for the winner and zero for all other outputs.\footnote{Technically, the winner output could be scaled to any magnitude.
We use $1$ for convenience.}
\todo{Have you tried decoding a threshold (say $\Theta(x_i - 0.15)$) off the LCA network?}
The LCA dynamics, however, prescribe an output of $u$ for the winner with $u$ being the largest input.
We define the `winner error' $E_1$ as the difference of the winner output from $u$ for the LCA network, as the difference from one for the IA network and, for both models, the `runner-up error' $E_2$ as the difference of second largest output from zero (with negative values treated as zero).
We use the average output values of the time   interval $(\SI{1}{\second}, \SI{2}{\second}]$ to calculate these error values (since these values fluctuate over time).
These two metrics only look at the final output averaged over time, but the model can produce transient outputs before the final decision is achieved.
Within the context of a larger model, this can be a problem as the transient output could already be interpreted as a decision.
Thus, we define $E_{2,\max}$ as the `highest output of a losing choice' during the whole simulation.

\section{Results}
We find that the ability to reach a clear decision depends on the magnitude of the highest input for both networks.
This is not surprising.
In the LCA network the output magnitude will be determined by the input magnitude of the largest input.
As such the largest input needs to be above 0.15 to fulfil the requirements.
For the IA network, the input has to be large enough to reach the integration threshold within the time limit.
By increasing the input scaling $\lambda$ or the time limit one could get a clear decision for smaller inputs within the time limit.
We further find that the IA network is independent of the input noise and target separation for all tested parameters.
This is not the case for the LCA network where the fraction of trials without a clear decision increases with noise and smaller target separation (see Fig.~\ref{fig:decisions}).
\begin{figure}
    \centering
    \includegraphics{figures/decisions}
    \caption{ \label{fig:decisions}
        Fraction of trials with a clear decision for different input noise standard deviations $\sigma$ and target separations $s$ for the LCA network.
        Error bars denote bootstrapped 95\% confidence intervals.
        The grey horizontal line shows the optimum.
    }
\end{figure}

However, when considering the correctness of the decisions (see Fig.~\ref{fig:correct}) the LCA network is doing better.
We can alleviate that difference to some degree by reducing the input scaling to $\lambda = 0.2$, but perform still worse for some combinations of input noise and target separation when staying below the time limit.
This in some part due to a systematic bias for certain choices introduced in the second layer of the IA network.
In the NEF the neuron gain and bias parameters are usually picked from random distributions.
As such the layer 2 populations will become active at slightly different thresholds.
By using identical layer 2 populations, i.e.~identical gain and bias parameters for each population, this bias can be eliminated and the IA network performs better than the LCA network in most instances.
\begin{figure*}
    \centering
    \includegraphics{figures/correct}
    \caption{ \label{fig:correct}
        Fraction of correct trials for different input noise standard deviations $\sigma$, target separations $s$, and models.  
        Error bars denote bootstrapped 95\% confidence intervals.
        The grey horizontal line shows the optimum.
    }
\end{figure*}

The time required to reach a decision in the LCA network depends mostly on the target separation and amount of noise on the input (see Fig.~\ref{fig:time}), the magnitude of the largest input is of minor influence.
Also, the confidence intervals indicate a larger variance in the decision times with less target separation and more input noise.
In contrast to that, in the IA network the scaled magnitude of the largest input is the most important factor.
Depending on this magnitude the network can either be faster or slower than the LCA network, but it will need more time to achieve the same fraction of correct responses.
There is also a slight influence of target separation and input noise with an interaction of these two parameters.
\begin{figure*}
    \centering
    \includegraphics{figures/time}
    \caption{ \label{fig:time}
        Mean decision times for different input noise standard deviations $\sigma$ and target separations $s$ for the LCA and IA networks.
        For the IA network $u = 0.2$ is shown which shows the interaction of target separation and input noise most clearly.
        Error bars denote bootstrapped 95\% confidence intervals.
        The grey horizontal lines show the optimum.
    }
\end{figure*}

The output of the IA network matches almost perfectly the specifications of producing an output of one for the winner and zero for all other choices.
This is not the case for the LCA network (see Fig.~\ref{fig:error}).
For small target separations the output of the winner will stay below $u$ and this will get worse with increased noise.
Furthermore, the LCA network will have residual outputs of losing choices under noisy conditions and with small target separations.
\begin{figure*}
    \centering
    \includegraphics{figures/error}
    \caption{ \label{fig:error}
        Error on outputs for different input noise standard deviations $\sigma$ and target separations $s$ for the LCA network.
        Error bars denote bootstrapped 95\% confidence intervals.
        The grey horizontal lines show the optimum.
    }
\end{figure*}

Finally, looking at the transient responses both models might produce outputs of losing choices (see Fig.~\ref{fig:transient}).
For the LCA network these increase with noise on the input and, for small target separation, can approach the final output of the winning choices.
For the IA network, transient outputs are smaller in noisy conditions, but can be higher than for the LCA network in less noisy conditions.
The magnitude of such transient responses is reduced by decreasing the input scaling.
\begin{figure*}
    \centering
    \includegraphics{figures/transient}
    \caption{ \label{fig:transient}
        Highest output of losing choices for different input noise standard deviations $\sigma$ and target separations $s$ for the LCA and IA networks.
        Error bars denote bootstrapped 95\% confidence intervals.
        The grey horizontal lines show the optimum.
    }
\end{figure*}
